Investigating Determinism of Game Engines for Autonomous Vehicle Verification

### 2020-04-18 paper/cav_determinisim_main.tex
gone fully through and added lots of todo notes, suggested restructuring to move case study before the derivation of the general methodology

@@

The determinism threshold depends on the objectives of the simulation and the granularity at which the simulation environment operates. In particular, this threshold must be sufficiently small to enable accurate assertion checking. \todo[inline]{Make sure assertion checking has previously been introduced.}
%

\todo[inline]{KIE: Move this to the Case Study section.}
In our case study, presented in Section~\ref{s:case-study}, a threshold must be chosen for our given application which reflects the scale at which repeatability of simulation execution is required. This value must be sufficiently small to allow accurate \textit{assertion checking} of the ego vehicle in the environment but a larger value may be acceptable for non-urban environments, e.g. motorway or highway. Similarly, the threshold value should not be so small as to fail with minor computational bottlenecks, e.g. $1\mu m$. Two autonomous vehicle projects this group are involved in, Robopilot and Capri, are looking at urban mobility and last mile delivery solutions which are predominantly concerned with urban environments rather than highway or motorway scenarios. Because of this, a deterministic threshold of $1m$ is insufficient when considering potentially dangerous close passing of cyclists and pedestrians and therefore a threshold of $1cm$ is more suitable.

@@

\subsection{Scenario Description}
The experiment should be a short scenario that can be easily repeated with the same initial conditions. Data logging should be used to record the actor positions at fixed time intervals throughout the simulation in order to determine the variance in actor path. The maximum deviation in actor path with respect to the mean for $n$-repeated tests will indicate at what dimensional scale deterministic results are achieved. For example, if 100 tests are executed and the maximum deviation in actor path is $<1cm$, then this scenario could be interpreted as being deterministic above this scale.

Tests should be sufficiently repeated using statistical power~\cite{cohen2013statistical} to reveal any outlier errors that occur with low probability. Initial testing\footnote{\label{note1}Experiments using the Unreal Engine not reported here for brevity but code and experimental results can be found at \url{https://github.com/TSL-UOB/AV-Determinism/tree/master/UnrealEngineTests}} indicated an actor path deviation of $1\times10^{-13} cm$ for 997 out of 1000 tests but with 3 tests reporting a deviation of over {\raise.17ex\hbox{$\scriptstyle\sim$}}$10cm$. While executing 100 repeats may seem sufficient, in this case the sample size may fail to observe these events that occur with low probability, inferring a false confidence in the results. To ensure outlier data points are not hidden in the average of a large number of tests, the maximum deviation should be reported rather than the average when summarising.

Game engine actors can be used to simulate the AV and other vehicles and pedestrians. For the experiment, actors should follow a series of trajectory waypoints which are hard-coded for each test type. In our experiments, hard-coded target location goals regularly spaced at $1m$ intervals were used for pedestrians and vehicles ensuring all actors were given the same initial conditions and any variation comes from execution of the simulator at runtime. Initial testing also indicated that when actors collide there is a significant change in \textit{simulation variance}, therefore the experiment should include some scenarios where collision callbacks are activated.

It is also important to control any pseudo-random numbers that are called in any external during the simulation with fixed seeds to ensure repeatable testing, e.g. external vehicle control programs or other clients connected to the simulator.

@@

\subsection{System Resources}
To simulate the demands of taxing simulations, potentially requiring realistic rendering or complex physics calculations in an otherwise simple scenario, stress programs should be used to artificially occupy system resources. Preliminary work\footnote{see footnote 14} suggested that adding stress to the CPU or GPU will  contribute towards increased \textit{simulation variance}. CPU utilisation software, such as the linux \texttt{stress} tool which is a workload generator program that can be used spawn workers on any number of cores or virtual threads on a system. This can be used to artificially increased the load on the system to explore at what point your system may become unstable leading to \textit{simulation variance}. For GPU utilisation \texttt{gpu-burn} was employed using the fur test at different resolutions and multiple instances can be used to generate high utilisation levels~\cite{GPU_stress}.

\subsection{Internal settings}
Where possible a fixed $dt$ should be used to give the most reliable physics calculations. If using Carla a small fixed value, say $0.05s$, can be set by using the following: \texttt{setting.fixed\_delta\_seconds = 0.05}, whereas in Unity the default fixed time step is set to $0.02s$\footnote{\url{https://docs.unity3d.com/ScriptReference/MonoBehaviour.FixedUpdate.html}} and a good description of the interplay between physics and render clock is given here~\cite{JohnAustinUnity}. In Carla, further settings can be applied to allow synchronous communication to external controllers to ensure no sensor data are passed out of order to the simulator~\cite{carla_sim_config}.

Within Unreal Engine there are numerous internal settings relating to the movement and interaction of physical bodies in the simulation. Settings can be adjusted to alter how actors interact and path plan via the navigation mesh of the environment, e.g. \textit{Contact Offset} and \textit{Navmesh Voxel Size}, or can be changed to improve the fidelity of physics calculations between game update steps, e.g. \textit{Physics Sub-Stepping} and \textit{Max Physics Delta Time}. Other options such as \textit{Enable Enhanced Determinism} were investigated along with running the engine from the command line with options for more deterministic behaviour \texttt{-deterministic}, floating point control \texttt{/fp:strict} and headless mode \texttt{-nullrhi} along with running the test as a packaged release by building and cooking\footnote{\url{https://docs.unrealengine.com/en-US/Engine/Deployment/Releasing/}}. However, none of these options improved simulation variance significantly beyond the default settings, see our previous investigation~\cite{TSLUnrealEngineTesting}.

@@

@@

\subsection{Output and Analysis}
The output for each experiment can be summarised as the \textit{maximum variance} of any actor over the $n-$repeated tests. The maximum variance is chosen to indicate the worst case rather than an average which could hide outlier data points when a high number of repeated tests are performed. This value can be analysed against the different scenarios or settings that have been identified as potential sources of simulation variance. These values can then be assessed against the user-defined \textit{deterministic threshold} to determine if the simulation is deterministic or not.

@@

% ***************************************************
%  CASE STUDY
% ***************************************************
\section{Case Study} \label{s:case-study}
We present a case study to illustrate some the challenges of simulation based verification outlined in Section~\ref{s:introduction}. The simulation must be deterministic and by our definition this should mean that the overall level of simulation variance $\sigma^2$ for any actor should be less than a user-defined threshold. Results are shown in terms of the maximum deviation ($\sigma$) and the deviation threshold in our case is $1cm$, where any value higher than this is considered non-deterministic.

In order to reproduce conditions for computationally intensive simulations, the experiments are executed with additional levels of artificial resource utilisation. Resource utilisation was artificially increased for both CPU and GPU devices (see Section \ref{s:methodology}) using reported values of the system monitors \texttt{htop} and \texttt{nvidia-smi} respectively. Resource utilisation figures reported here should be considered approximate values.
% The level of additional resource utilisation, above that of the needs of the simulation, was increased from $0\%$ (no additional resource load) in $25\%$ increments up to a maximum of $95\%$ for both CPU and GPU as reported by system monitors (\texttt{htop} and \texttt{nvidia-smi} respectively).

@@

% \subsection{Tests Description And Technicalities}\label{TestsDescriptionAndTechnicalities}
\subsection{Scenario Description}\label{TestsDescriptionAndTechnicalities}
The list of tests shown in Table \ref{TableOfExperiments} were chosen to cover a range of interactions between actor types. Tests 1 \& 2 involve 2 vehicles meeting at a junction where they do not collide (test 1) and when they do tiggering a collision callback in the game engine (test 2). In both cases the trajectories of the vehicles are hard-coded to follow a set of waypoints spaced at $0.1m$ intervals using a PID controller. In test 3 a mixture of different actor types are introduced where two vehicles dive without collision and a pedestrian walks across the road at a crossing point. In test 4 this pedestrian collides with one of the vehicles at the crossing triggering a collision callback, see Fig~\ref{Test_a}. Similar to vehicles, pedestrians navigate via a set of regularly space waypoints at $0.1m$ intervals using the A* algorithm which is the default controller for the Carla pedestrian actors. Tests 5 \& 6 involve only pedestrians where they do not collide (test 5) and where they do (test 6), see Fig.~\ref{Test_b}.

@@

% ***************************************************
%  RESULTS
% ***************************************************
\subsection{Results And Discussion}\label{ResultsSection}
The summary of all the tests are shown in the final column of Table~\ref{TableOfExperiments} where $\max\sigma$ is the maximum deviation across all resource utilisation levels. From these summary results it is clear that scenarios with only pedestrian actors are those that are consistently deterministic over all resource utilisation levels. However, for other scenarios involving vehicles or a mixture of actors types resource utilisation level has a significant impact on \textit{simulation variance}.

Figure~\ref{ExperimentsStressSummary} shows $\max\sigma$ against resource utilisation level, where the $x$-axis indicates the approximate percentage of resource utilisation (CPU \& GPU). In this figure, anything above the $cm$ threshold (dashed line) is considered non-deterministic according to our specific threshold level informed from our verification requirements. A general pattern in the results indicates that some scenarios fail to produce deterministic results consistently against any resource utilisation level and that some cases only fail at utilisation levels above 50\%.

However, the non-deterministic results in Figure~\ref{ExperimentsStressSummary} (all those above the dashed line) are the worst case account of the situation. Examining these specific results, tests 2 \& 4, in more detail as a function of simulation time reveals further information about the simulation variance before and after an actor collision. Fig.~\ref{CarsCollisionCG25} shows this examination for vehicle to vehicle collisions (test 2), where $\max\sigma$ switches from deterministic prior to the vehicle collision and is non-deterministic post-collision. This pattern of deterministic behaviour prior to collision is maintained up to a resource utilisation level of approximately 75\%, Fig.~\ref{CarsCollisionPrePost}. This time series examination was repeated for vehicle to pedestrian collisions (test 4) and the results are shown in Fig.~\ref{CarsPeopleCollsionsCG25}. Similarly to vehicle collisions, the variation of $\max\sigma$ for vehicle to pedestrian collisions indicates pre-collision deterministic behaviour with up to 75\% resource utilisation, Fig.~\ref{CarsPeopleCollisionPrePost}. This is a key finding of this work and suggests that verification engineers should consider this as an end condition for terminating tests as any post-collision results are not reliable.

The second key finding of this work is illustrated in Fig.~\ref{CarsPeopleCollsionsCG25}. In this scenario (test 4), there is a collision between Car ID 2 (solid line) and Ped ID 3 (dot dash line) which occurs at a simulation time of approximately $6s$ and the other vehicle, Car ID 1, is \textit{not involved in the collision}. There are three key observations here; firstly that the vehicle directly involved in the collision (Car ID 2) displays non-deterministic actor position post-collision as describe above. Secondly, that the pedestrian involved in the collision (Ped ID 3) displays deterministic behaviour throughout the test\footnote{note that in Carla the pedestrian object is destroyed post-collision hence the flat line following the collision}. And, thirdly we observed a delayed effect on Car ID 1 showing high simulation variance with a $5s$ delayed effect \textit{even though this vehicle was not involved in the collision}. This final point should be of particular concern to verification engineers and research practitioners in the field as it implies that any collision between actors will affect the simulation variance of all actors in the simulation and could potentially result in false-positive assertion violations.

@@

\textbf{Big question why does high-RU influence simulation variance?? Can we tie this to thread scheduling and will NUMAclt change this conclusion?'}

@@

@@

% ***************************************************
%  CONCLUSION
% ***************************************************
\section{Conclusions}\label{s:conclusion}
Practitioners in this field can resolve the non-deterministic operational domains of their simulation by following the method outlined in this paper. Potential sources of non-deterministic simulation behaviour have been discussed such as floating point rounding error and scheduling policy. A case study was used to ascertain the boundary between deterministic and non-deterministic operation of the Carla driving simulator. Results suggest non-deterministic outcomes when resource utilisation levels exceeded 75\% or when vehicle collisions occur for the particular hardware used in this study. Furthermore it was found that all actors, even those not involved directly in a collision, will behave non-deterministically if a collision callback to the game engine is made. Best operating practice follows that the simulation should be stopped as soon as a collision callback is detected or resource utilisation levels exceed 75\% although this value may be system specific.

>>>>

\todo[inline]{We may need to introduce the Case Study first, then derive the general Methodology from it. This way, we can explain the requirements of the case study, how things were set up, what was observed, how the methodology was refined, several iterations even, etc.}

@@

The determinism threshold depends on the objectives of the simulation and the granularity at which the simulation environment operates. In particular, this threshold must be sufficiently small to enable accurate assertion checking. \todo{Make sure assertion checking has previously been introduced.}
%
In practice, it may be necessary to determine this threshold experimentally.

\todo[inline]{KIE: Move this to the Case Study section and justify the values.
In our case study, presented in Section~\ref{s:case-study}, a threshold must be chosen for our given application which reflects the scale at which repeatability of simulation execution is required. This value must be sufficiently small to allow accurate \textit{assertion checking} of the ego vehicle in the environment but a larger value may be acceptable for non-urban environments, e.g. motorway or highway. Similarly, the threshold value should not be so small as to fail with minor computational bottlenecks, e.g. $1\mu m$.}
%
\todo[inline]{Move to Case Study, rewrite with justification: Two autonomous vehicle projects this group are involved in, Robopilot and Capri, are looking at urban mobility and last mile delivery solutions which are predominantly concerned with urban environments rather than highway or motorway scenarios. Because of this, a deterministic threshold of $1m$ is insufficient when considering potentially dangerous close passing of cyclists and pedestrians and therefore a threshold of $1cm$ is more suitable.}\todo{Formal writing, no reference to the group or project, please.}

@@

\subsection{Scenario Description}\todo[inline]{Select section titles to match the flow diagram. In each of these sections, go through the matching phase in the flow diagram, explaining each point listed there and justify recommendations.}
The experiment should be a short\todo{What does short mean? Quantify.} scenario that can be easily repeated with the same initial conditions. Data logging should be used to record the actor positions at fixed time intervals throughout the simulation in order to determine the variance in actor path. The maximum deviation in actor path with respect to the mean for $n$ repeated tests will indicate at what dimensional scale deterministic results are achieved. For example, if 100 tests are executed and the maximum deviation in actor path is $<1cm$, then this scenario could be interpreted as being deterministic above this scale.

Tests should be sufficiently repeated using statistical power~\cite{cohen2013statistical} to reveal any outlier errors that occur with low probability.\todo{Explain what this means.} Initial testing\footnote{\label{note1}Experiments using the Unreal Engine not reported here for brevity but code and experimental results can be found at \url{https://github.com/TSL-UOB/AV-Determinism/tree/master/UnrealEngineTests}} indicated an actor path deviation of $1\times10^{-13} cm$ for 997 out of 1000 tests but with 3 tests reporting a deviation of over {\raise.17ex\hbox{$\scriptstyle\sim$}}$10cm$. While executing 100 repeats may seem sufficient, in this case the sample size may fail to observe these events that occur with low probability, inferring a false confidence in the results. To ensure outlier data points are not hidden in the average of a large number of tests, the maximum deviation should be reported rather than the average when summarising.

Game engine actors can be used to simulate the AV and other vehicles as well as pedestrians. For the experiment, actors should follow a series of trajectory waypoints which are hard-coded for each test type.

\todo[inline]{This is mixing our experiments (which are part of the case study) with the methodology. It is probably best to first introduce the case study and then derive the methodology from it, rather than mixing both.} In our experiments, hard-coded target location goals regularly spaced at $1m$ intervals were used for pedestrians and vehicles ensuring all actors were given the same initial conditions and any variation comes from execution of the simulator at runtime. Initial testing also indicated that when actors collide there is a significant change in \textit{simulation variance}, therefore the experiment should include some scenarios where collision callbacks are activated.\todo[inline]{Make sure collision callbacks have been introduced earlier.}

Randomisation must be fully controlled by fixing the seeds for any random number generation that is used during the experiment. It is important to identify and control any external sources that may use randomisation, e.g.\ external vehicle controllers or other clients connected to the simulation environment.

@@

\subsection{Internal settings}\todo[inline]{Systematically go through the points listed in the flow diagram, explain and justify.}
Where possible a fixed $dt$ should be used to give the most reliable physics calculations. If using Carla a small fixed value, say $0.05s$, can be set by using the following: \texttt{setting.fixed\_delta\_seconds = 0.05}, whereas in Unity the default fixed time step is set to $0.02s$\footnote{\url{https://docs.unity3d.com/ScriptReference/MonoBehaviour.FixedUpdate.html}} and a good description of the interplay between physics and render clock is given here~\cite{JohnAustinUnity}. In Carla, further settings can be applied to allow synchronous communication to external controllers to ensure no sensor data are passed out of order to the simulator~\cite{carla_sim_config}.

Within Unreal Engine there are numerous internal settings relating to the movement and interaction of physical bodies in the simulation. Settings can be adjusted to alter how actors interact and path plan via the navigation mesh of the environment, e.g. \textit{Contact Offset} and \textit{Navmesh Voxel Size}, or can be changed to improve the fidelity of physics calculations between game update steps, e.g. \textit{Physics Sub-Stepping} and \textit{Max Physics Delta Time}.

Other options such as \textit{Enable Enhanced Determinism} were investigated along with running the engine from the command line with options for more deterministic behaviour \texttt{-deterministic}, floating point control \texttt{/fp:strict} and headless mode \texttt{-nullrhi} along with running the test as a packaged release by building and cooking\footnote{\url{https://docs.unrealengine.com/en-US/Engine/Deployment/Releasing/}}. However, none of these options improved simulation variance significantly beyond the default settings, see our previous investigation\footnote{\url{https://github.com/TSL-UOB/AV-Determinism/tree/master/UnrealEngineTests}}.

@@

\subsection{System Resources}\todo[inline]{This should be ``External Settings'' - from flow diagram?}
To simulate the demands of taxing simulations, potentially requiring realistic rendering or complex physics calculations in an otherwise simple scenario, stress programs should be used to artificially occupy system resources. Preliminary work\footnote{see footnote 14} suggested that adding stress to the CPU or GPU will  contribute towards increased \textit{simulation variance}. CPU utilisation software, such as the linux \texttt{stress} tool which is a workload generator program that can be used spawn workers on any number of cores or virtual threads on a system. This can be used to artificially increased the load on the system to explore at what point your system may become unstable leading to \textit{simulation variance}. For GPU utilisation \texttt{gpu-burn} was employed using the fur test at different resolutions and multiple instances can be used to generate high utilisation levels~\cite{GPU_stress}.\todo[inline]{Again, this sounds as if having the case study first makes more sense. That way we can explain wrt the experiments and then generalise to obtain the Methodology.}

@@

\subsection{Output and Analysis}
\todo[inline]{Say what needs to be logged, what needs to be calculated based on these logs and what one should be looking for and why.}
The output for each experiment can be summarised as the \textit{maximum variance} of any actor over the $n$ repeated tests. The maximum variance is chosen to indicate the worst case rather than an average which could hide outlier data points when a high number of repeated tests are performed. This value can be analysed against the different scenarios or settings that have been identified as potential sources of simulation variance. These values can then be assessed against the user-defined \textit{deterministic threshold} to determine if the simulation is deterministic or not.

@@

% ***************************************************
%  CASE STUDY
% ***************************************************
\section{Case Study} \label{s:case-study}\todo[inline]{Put this section before Methodology.}
We present a case study to illustrate some the challenges of simulation based verification outlined in Section~\ref{s:introduction}. The simulation must be deterministic and by our definition this should mean that the overall level of simulation variance $\sigma^2$ for any actor should be less than a user-defined threshold. Results are shown in terms of the maximum deviation ($\sigma$) and the deviation threshold in our case is $1cm$, where any value higher than this is considered non-deterministic.

\todo[inline]{Introduce the Scenario - see my comment later.}

\todo[inline]{Set the threshold, explain what a good threshold is, available options and what was selected and why.}

\todo[inline]{Again, match the flow diagram steps in the subsections, but here wrt the case study steps.}

\subsection{External Settings}\todo[inline]{This would be ``External Settings'' and can be introduced after finding that internal settings alone did not tame the non-determinism. :(}
In order to reproduce conditions for computationally intensive simulations, the experiments are executed with additional levels of artificial resource utilisation. Resource utilisation was artificially increased for both CPU and GPU devices (see Section \ref{s:methodology}) using reported values of the system monitors \texttt{htop} and \texttt{nvidia-smi} respectively. Resource utilisation figures reported here should be considered approximate values.
% The level of additional resource utilisation, above that of the needs of the simulation, was increased from $0\%$ (no additional resource load) in $25\%$ increments up to a maximum of $95\%$ for both CPU and GPU as reported by system monitors (\texttt{htop} and \texttt{nvidia-smi} respectively).

@@

% \subsection{Tests Description And Technicalities}\label{TestsDescriptionAndTechnicalities}
\subsection{Scenario Description}\label{TestsDescriptionAndTechnicalities}\todo[inline]{This should go first.}
\todo[inline]{First describe the scenario(s), WHAT happens, HOW does it happen and WHY is this scenario interesting for this investigation, i.e.\ what are the features this scenario has? Then say how this has been implemented, referring to test numbers and the }
The tests listed in Table~\ref{TableOfExperiments} were chosen to cover a range of interactions between actor types. Tests 1 \& 2 involve 2 vehicles meeting at a junction where they do not collide (test 1) and when they do tiggering a collision callback in the game engine (test 2). In both cases the trajectories of the vehicles are hard-coded to follow a set of waypoints spaced at $0.1m$ intervals using a PID controller. In test 3 a mixture of different actor types are introduced where two vehicles drive without collision and a pedestrian walks across the road at a crossing point. In test 4 this pedestrian collides with one of the vehicles at the crossing, triggering a collision callback, see Fig~\ref{Test_a}. Similar to vehicles, pedestrians navigate via a set of regularly spaced waypoints at $0.1m$ intervals using the A* algorithm which is the default controller for the Carla pedestrian actors. Tests 5 \& 6 involve only pedestrians that do not collide (test 5) and that do collide (test 6), see Fig.~\ref{Test_b}.

@@

% ***************************************************
%  RESULTS
% ***************************************************
\subsection{Results and Discussion}\label{ResultsSection}
The summary of all the tests are shown in the final column of Table~\ref{TableOfExperiments} where $\max\sigma$ is the maximum deviation across all resource utilisation levels. From these  results it is clear that scenarios with only pedestrian actors are those that are consistently deterministic over all resource utilisation levels. However, for other scenarios involving vehicles or a mixture of actors types, the resource utilisation level has a significant impact on \textit{simulation variance}.

Figure~\ref{ExperimentsStressSummary} shows $\max\sigma$ against the resource utilisation level, where the $x$-axis indicates the approximate percentage of resource utilisation (CPU \& GPU). In this figure, anything above the $1cm$ threshold (dashed line) is considered non-deterministic according to our specific threshold level informed from our verification requirements.\todo[inline]{Where are the verification requirements listed? They should be part of the scenario description.} A general pattern in the results indicates that some scenarios fail to produce deterministic results consistently against any resource utilisation level and that some cases only fail at utilisation levels above 50\%.\todo[inline]{Help the reader see what you can see by explaining exactly where this can be seen in the figure(s).}

However, the non-deterministic results in Figure~\ref{ExperimentsStressSummary} (all those above the dashed line) are the worst case account of the situation. Examining these specific results, tests 2 \& 4, in more detail as a function of simulation time reveals further information about the simulation variance before and after an actor collision. Fig.~\ref{CarsCollisionCG25} shows this examination for vehicle to vehicle collisions (test 2), where $\max\sigma$ switches from deterministic prior to the vehicle collision to non-deterministic post-collision. This pattern of deterministic behaviour prior to collision is maintained up to a resource utilisation level of approximately 75\%, see Fig.~\ref{CarsCollisionPrePost}. This time series examination was repeated for vehicle to pedestrian collisions (test 4) and the results are shown in Fig.~\ref{CarsPeopleCollsionsCG25}. Similarly to vehicle collisions, the variation of $\max\sigma$ for vehicle to pedestrian collisions indicates deterministic pre-collision behaviour with up to 75\% resource utilisation, see Fig.~\ref{CarsPeopleCollisionPrePost}. This is a key finding; it suggests that verification engineers should consider terminating tests at the point of a collision, as any post-collision results cannot be relied upon due to the observed non-determinism. \todo{Why ``not reliable''.}

The second key finding of this work is illustrated in Fig.~\ref{CarsPeopleCollsionsCG25}. In this scenario (test 4), there is a collision between Car ID 2 (solid line) and Ped ID 3 (dot dash line) which occurs at a simulation time of approximately $6s$ and the other vehicle, Car ID 1, is \textit{not involved in the collision}. There are three key observations here; firstly that the vehicle directly involved in the collision (Car ID 2) displays non-deterministic actor position post-collision as described above. Secondly, that the pedestrian involved in the collision (Ped ID 3) displays deterministic behaviour throughout the test\footnote{Note that in Carla the pedestrian object is destroyed post-collision hence the flat line following the collision.}. Thirdly, we observed a delayed effect on Car ID 1 showing high simulation variance with a $5s$ delay \textit{even though this vehicle was not involved in the collision}. This final point should be of particular concern to verification engineers and research practitioners in the field as it implies that any collision between actors can affect the simulation variance of all actors in the simulation and could potentially result in  false-positive assertion violations.\todo[inline]{Make sure that assertion checking has been introduced before.}

@@

\todo[inline]{Big question why does high-RU influence simulation variance?? Can we tie this to thread scheduling and will NUMAclt change this conclusion?'}

@@

\todo[inline]{Sum this up and lead on to the Methodology - assuming that this section is placed before the Methodology section. This should cover that our results are clearly specific to our HW and provide evidence of non-determinism, but the results are not transferable. Thus, we present a general Methodology that others can re-use/apply to determine whether or not their sim is deterministic - in the next section.}

@@

% ***************************************************
%  CONCLUSION
% ***************************************************
\section{Conclusions and Future Work}\label{s:conclusion}

\todo[inline]{Say what the paper investigated and what we aimed to achieve, why this is important and how we went about it. Then point out the derivation of a re-usable methodology based on the knowledge and understanding we gained from the case study.}

\todo[inline]{Review/rewrite this to match the earlier changes in the paper. Practitioners in this field can resolve the non-deterministic operational domains of their simulation by following the method outlined in this paper. Potential sources of non-deterministic simulation behaviour have been discussed such as floating point rounding error and scheduling policy. A case study was used to ascertain the boundary between deterministic and non-deterministic operation of the Carla driving simulator. Results suggest non-deterministic outcomes when resource utilisation levels exceeded 75\% or when vehicle collisions occur for the particular hardware used in this study. Furthermore it was found that all actors, even those not involved directly in a collision, will behave non-deterministically if a collision callback to the game engine is made. Best operating practice follows that the simulation should be stopped as soon as a collision callback is detected or resource utilisation levels exceed 75\% although this value may be system specific. }

\todo[inline]{Point out (again) that all data and scripts etc are on github.}

\todo[inline]{Future work...}
